Visualization and analysis of probability distributions of large temporal data

### 2022-02-09 report/response.Rmd

I thank my examiners for their thorough and constructive comments. The point by point description of changes are below: the examiners' comments are in red and my response is in black.

# Professor Catherine Hurley

I appreciate all the comments and suggestions, and will incorporate them into future work. Specific comments are addressed as follows:

## Chapter 1

Perhaps though, this would be a good place to detail the co-authorship arrangements for the succeeding chapters, and for the student’s contribution to be stated clearly.

@@

> Q2. Last sentence of Section 2.3.4: fix the table reference

Done.

> Q3. In Figure 2.4, the outlier glyphs are in my opinion too big and dark and attract too much attention.
The log scale for the y-axis should be mentioned in the caption.

Done. 
The outliers are made smaller and transparent and caption changed.

@@

Not change.
Due to the space constraint in the paper.

## Chapter 3

> Q1. On page 33 there is a mention of the hakear package, without a reference.

No change.
It appears in the abstract of the paper. For the thesis, I have added it since the abstract acts as a summary to the chapter. The package reference is mentioned in the Supplementary section of the paper. Moreover, the plan is to bring all the functions from `gravitas` to `hakear` when we publish it.

@@

> Q11. Add v to table 3.1.

Done.
Added v:continuous univariate measured variable

> Q12. In the first sentence on the top of page 42, ordered and unordered are mixed up. Have you considered
the setting where the facet variable levels are ordered? For the within-facet ordered distances, you
could consider a distance measure that respects circular order, or choose the start level appropriate to
the display.

first part: Done.
second part: This is a good suggestion. We will include that into future work.

@@

CHECK.
No change.
We assume we have enough observations for each level or combination of levels to compute distributions. As long as that holds, no adjustment made for varying number of observations.

> Q15. The equations on page 44 could be tidied up. There are extra () on the definition of wpdperm and on the residual definition at the bottom of the page. Use $\times$ for the equations.

Done.

> Q16. Is there a practical reason why wpdglm is not working as expected for lower nx and nf?

CHECK.
No practical reason but a more theoretical one.
For lower $n_x$ and $n_f$, the number of combinations $n_xn_f$ is low such that the max do not have an assymptotic distribution which can be assumed as Normal by CLT. With higher combinations, the distribution of the each combination could be assumed to be normal.

@@

> Q17. In the 3.3.1 algorithm, there is m, and then M.

Changed.
1.b) for $i \in {1, 2, \dots, M}$.

> Q18. In Table 3.2, I would suggest separating the m = 1 and m = 2 tables.

Not changed.

> Q19. The presentation of the material on the simulation study on page 48 could be improved. Where is the notation $wpd_{l,s}$ used? Figure 3.5 shows the m = 2 results only. The text alludes to a simulation involving different underlying distributions, but this is not mentioned again.

DISCUSS
first part: How?
second part: After NQT, results are pretty similar to an underlying normal distribution. (Discuss)

@@

 $\mu$ is the difference between means considered for consecutive categories.

> Q11. In Section 4.3.3 what are the values of R and T?

> Q12. Figure 4.5 caption: MDS summary plots of what?

MDS plot summary of the data sets from the validation study.

@@

> Q13. Figure 4.7 is missing labels (a), (b) (c).

Done. Added a, b representing 12 customers (rows) in each of them.

@@

> Q14. Why do you chose to summarize 4 and 5 clusters from JS-NQT when sindex suggests 3 clusters?

It is told in the "number of clusters" section that the number of optimal number of clusters is a difficult exercise. The idea was to show how decreasing the number of clusters can hide distinct patterns.

@@

Done.

> Q17. Again, the computational burden could do with more Discussion. At present, if I were to use this
method on my data, what kind of sizes are realistic to work with?

# Prof Juergen Symanzik

## Chapter 1

> Q1. While the author hints at other sources of similar data, I think it would be worthwhile to give the reader some further specific ideas here in the Introduction where similar complex data may occur, e.g., data from traffic sensors, movement data (turnstiles in metro stations or public buildings), and even more traditional measurements such as temperatures or precipitation over
time that could benefit from the methods and software described in this thesis.

> Q2. It may also be helpful to state what is not be covered in this thesis: The spatial component of such data where an additional component could be latitude and longitude or some areal code,
such as the post code or some administrative units. Extending the work from this thesis with spatial proximity methods for clustering (https://doi.org/10.1016/j.cageo.2011.12.017) [assuming that households in certain residential neighborhoods may have similar energy usage patterns] and trying spatial visualization methods (if possible) such as glyph maps (https:
//onlinelibrary.wiley.com/doi/abs/10.1002/env.2152) should be mentioned somewhere as a consideration for future work | but that may become a future new thesis by itself.

> Q3. Finally, it would be helpful to indicate here (and in each chapter) what the names or abbreviations for the three R packages (gravitas, hakear, and gracsr) represent. Even the full R package
title of "gravitas: Explore Probability Distributions for Bivariate Temporal Granularities" does not provide a full answer what the "tas" component means (I am assuming that Gravi stands for Granularities Visualization). I am even more lost with hakear.

> Q1. One reference to a table (Table 2.3) does not resolve and appears as @ref{tab:tab-mayan} in the main text on p. 16.

> Q2. how to define and handle pay day (or loan day) which often is defined as the 1st (and/or possibly 15th) work day of a month, resulting in some aperiodic granularity as some of these days across the year could initially fall on a weekend, but even more in a few
cases, may coincide with a public holiday (and thus shifting pay day by 2 or 3 days). Similarly, extended weekends that could stretch from Fridays to Sundays or from Saturdays to Mondays (and possibly even over a 4{day period) might be worth some discussion, in particular as a case of variable{length aperiodic granularity. This latter granularity may be of particular interest when working with temporal data with an economic (travel, restaurants, etc.) or entertainment (movie theaters, casinos, etc.) aspect.

@@

> Q9. It would be helpful to add a specific link to a subsection in a cross-reference such as (See the supplements for more details.) on p. 42. Same on p. 47 and p. 48.

Done. (Adding link in supplementary section enough?)

> Q10. On p. 45 and in (3.2), it is not immediately clear whether nx; nf simultaneously have to be less than or equal to 5 so that wpdperm is being used - or whether only one of them has to be less than 5. Just a verbal clarification is needed. A practical scenario where this applies might be months (12) and weekdays/weekend (2).

Done.
We have assumed both nx; nf simultaneously to be less than or equal to 5

### 2022-02-09 Rmd/03-hakear.Rmd
Ch 4 Q1 Juergen

These results justify our use of the permutation approach when $\nx<5$ and $\nf<5$, and the use of the GLM otherwise.
<!--
## Choosing the threshold

@@

According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least. From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an inconsequential pair for id 1. The distribution of energy demand for both of these households, with *dow* as the facet and *hod* on the x-axis, may help explain the choice. Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates for most hours of the day and days of the week, while for id 1, daily patterns are more consistent within weekdays and weekends. As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wdwnd*), which has been rated higher in Table \ref{tab:tab-rank-8}.

@@

(ref:gravitas-plot-8) Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) across *hod* in x-axis and *dow* in facets through quantile area plots. The value of $\wpd$ in Table 3 suggests that the harmony pair (*dow*, *hod*) is significant for household id 7, but not for id 1. This implies that distributional differences are captured more by this harmony for id 7, which is apparent from this display with more fluctuations across median and 75th percentile for different hours of the day and day of week. For id 1, patterns look similar within weekdays and weekends. Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.

>>>>

These results justify our use of the permutation approach when $\nx \leq 5$ and $\nf \leq 5$, and the use of the GLM otherwise.
<!--
## Choosing the threshold

@@

According to Figure \ref{fig:dotplot-8}(c), for the harmony pair (*dow*-*hod*), household id 7 has the greatest value of $\wpd$, while id 1 has the least. From Table \ref{tab:tab-rank-8} it can be seen that the harmony pair (*dow*, *hod*) is important for id 7; however, it has been labeled as an insignificant pair for id 1. The distribution of energy demand for both of these households, with *dow* as the facet and *hod* on the x-axis, may help explain the choice. Figure \ref{fig:gravitas-plot-8} demonstrates that for id 7, the median (black) and quartile deviation (orange) of energy consumption fluctuates for most hours of the day and days of the week, while for id 1, daily patterns are more consistent within weekdays and weekends. As a result, for id 1, it is more appropriate to examine the distributional difference solely across (*dow*, *wdwnd*), which has been rated higher in Table \ref{tab:tab-rank-8}.

@@

(ref:gravitas-plot-8) Comparing distribution of energy demand shown for household id 1 (a) and 7 (b) on logarithmic scale across *hod* in x-axis and *dow* in facets through quantile area plots. The value of $\wpd$ in Table 3 suggests that the harmony pair (*dow*, *hod*) is significant for household id 7, but not for id 1. This implies that distributional differences are captured more by this harmony for id 7, which is apparent from this display with more fluctuations across median and 75th percentile for different hours of the day and day of week. For id 1, patterns look similar within weekdays and weekends. Here, the median is represented by the black line, the orange area corresponds to quartile deviation and the green area corresponds to area between $10^{th}$ and $90^{th}$ quantile.
